---
title: "Show the right numbers, 1"
linktitle: "08: Show the right numbers"
date: "2020-09-11"
menu:
  show:
    parent: Examples
    weight: 1
type: docs
bibliography: "../../static/bib/references.bib"
csl: "../../static/bib/chicago-fullnote-bibliography-no-bib.csl"
output:
  blogdown::html_page:
    toc: false
editor_options: 
  chunk_output_type: console
---

```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
``` 

```{r plot1, message = FALSE}
library(tidyverse)
library(gapminder)
library(socviz)
```

### Geoms can transform data

```{r bar1}

p <- ggplot(data = gss_sm,
            mapping = aes(x = bigregion))
p + geom_bar()

```

```{r bar2}

p <- ggplot(data = gss_sm,
            mapping = aes(x = bigregion))
p + geom_bar(mapping = aes(y = ..prop..))

```

```{r bar3}

p <- ggplot(data = gss_sm,
            mapping = aes(x = bigregion))
p + geom_bar(mapping = aes(y = ..prop.., group = 1)) 


```


### Histograms and Density Plots 

```{r hist_1}

p <- ggplot(data = midwest,
            mapping = aes(x = area))
p + geom_histogram()

p <- ggplot(data = midwest,
            mapping = aes(x = area))
p + geom_histogram(bins = 10)

```


```{r hist_2}

oh_wi <- c("OH", "WI")

p <- ggplot(data = subset(midwest, subset = state %in% oh_wi),
            mapping = aes(x = percollege, fill = state))
p + geom_histogram(alpha = 0.4, bins = 20)

```


```{r density_1}

p <- ggplot(data = midwest,
            mapping = aes(x = area))
p + geom_density()


```

```{r density_2}


p <- ggplot(data = midwest,
            mapping = aes(x = area, fill = state, color = state))
p + geom_density(alpha = 0.3)

```

### Avoid transformations when necessary


```{r titanic_1}

titanic

```

```{r titanic_2}

p <- ggplot(data = titanic,
            mapping = aes(x = fate, y = percent, fill = sex))
p + geom_bar(position = "dodge", stat = "identity") + theme(legend.position = "top")

```


```{r lifegap_1}

oecd_sum

```

```{r lifegap_2, fig.height = 4, fig.width = 9, layout = 'l-page'}

p <- ggplot(data = oecd_sum,
            mapping = aes(x = year, y = diff, fill = hi_lo))
p + geom_col() + guides(fill = FALSE) +
  labs(x = NULL, y = "Difference in Years",
       title = "The US Life Expectancy Gap",
       subtitle = "Difference between US and OECD average life expectancies, 1960-2015",
       caption = "Data: OECD. After a chart by Christopher Ingraham,
                  Washington Post, December 27th 2017.")

```

### Frequency Plots the Slightly Awkward Way

```{r gss_tab}

with(gss_sm, table(religion))

```

```{r gss_color_fill}


p <- ggplot(data = gss_sm,
            mapping = aes(x = religion, color = religion))
p + geom_bar()

p <- ggplot(data = gss_sm,
            mapping = aes(x = religion, fill = religion))
p + geom_bar() + guides(fill = FALSE)

```

```{r two_way_1}

p <- ggplot(data = gss_sm,
            mapping = aes(x = bigregion, 
                          fill = religion))
p + geom_bar()

```

```{r two_way_2}
p <- ggplot(data = gss_sm,
            mapping = aes(x = bigregion, 
                          fill = religion))
p + geom_bar(position = "fill")      
```

```{r two_way_3}
p <- ggplot(data = gss_sm,
            mapping = aes(x = bigregion, 
                          fill = religion))
p + geom_bar(position = "dodge",
             mapping = aes(y = ..prop..))      
```


```{r two_way_4}
p <- ggplot(data = gss_sm,
            mapping = aes(x = bigregion, 
                          fill = religion))
p + geom_bar(position = "dodge",
             mapping = aes(y = ..prop.., 
                           group = religion))       
```

This gets awkward and a bit confusing quite fast. In general it's better not to try to make ggplot do too much right at the plotting stage. Instead, pre-compute the numbers you want to show. We'll see how to do that next.


## Dplyr pipelines

```{r pipeline}

rel_by_region <- gss_sm %>%
    group_by(bigregion, religion) %>%
    tally() %>%
    mutate(freq = n / sum(n),
           pct = round((freq*100), 1))

```


```{r 05-tables-and-labels-1 }
rel_by_region
```
