---
title: "Show the right numbers, 2"
linktitle: "9: Show the right numbers, 2"
date: "2020-09-16"
menu:
  show:
    parent: Examples
    weight: 1
type: docs
bibliography: "../../static/bib/references.bib"
csl: "../../static/bib/chicago-fullnote-bibliography-no-bib.csl"
output:
  blogdown::html_page:
    toc: false
editor_options: 
  chunk_output_type: console
---

```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
``` 

```{r plot1, message = FALSE}
library(tidyverse)
library(gapminder)
library(socviz)
```

## A dplyr pipeline

```{r pipeline}

rel_by_region <- gss_sm %>%
    group_by(bigregion, religion) %>%
    tally() %>%
    mutate(freq = n / sum(n),
           pct = round((freq*100), 1))

```


```{r 05-tables-and-labels-1 }
rel_by_region
```

## n(), tally(), and count()

```{r 05-aside-n}
gss_sm %>%
    group_by(bigregion, religion) %>%
    summarize(n = n())
```

```{r 05-aside-tally}
gss_sm %>%
    group_by(bigregion, religion) %>%
    tally()

```

```{r 05-aside-count}
gss_sm %>%
    count(bigregion, religion)
```

## Table checking
```{r 05-tables-and-labels-2 }
rel_by_region %>% group_by(bigregion) %>%
    summarize(total = sum(pct))      
```


## Facets are better than fill mappings

```{r dodge2}

p <- ggplot(rel_by_region, aes(x = bigregion, y = pct, fill = religion))
p + geom_col(position = "dodge") +
    labs(x = "Region",y = "Percent", fill = "Religion") +
    theme(legend.position = "top")      
```

```{r facet_dodge, fig.height = 2.5, fig.width = 8}
p <- ggplot(rel_by_region, aes(x = pct, y = religion, fill = religion))
p + geom_col() +
    labs(x = "Percent", y = NULL, fill = "Religion") +
    guides(fill = FALSE) + 
    facet_wrap(~ bigregion, nrow = 1)      
```

## facet_wrap and facet_grid

```{r grid_v_wrap}
p <-  ggplot(data = gss_sm,
             mapping = aes(x = age, y = childs))


p + geom_point(alpha = 0.2) + geom_smooth() +
    facet_wrap(sex ~ race)

p + geom_point(alpha = 0.2) + geom_smooth() +
  facet_grid(sex ~ race)


p + geom_point(alpha = 0.2) + geom_smooth() +
  facet_grid(bigregion ~ sex + race)

```


## The Organ Donation Data

```{r organdata}
organdata
```

```{r 05-tables-and-labels-3 }
organdata %>% select(1:6) %>% sample_n(size = 10)      
```

```{r 05-tables-and-labels-4 }
p <- ggplot(data = organdata,
            mapping = aes(x = year, y = donors))
p + geom_point()      
```

```{r 05-tables-and-labels-5 }
p <- ggplot(data = organdata,
            mapping = aes(x = year, y = donors))
p + geom_line(mapping = aes(group = country))      
```


- Control the height and width of your figures with, for example, `fig.width = 12, fig.height = 8` in the options to the code chunk that draws the plot.

```{r grouped_and_faceted, fig.width = 12, fig.height = 8}
p <- ggplot(data = organdata,
            mapping = aes(x = year, y = donors))
p + geom_line() + 
  facet_wrap(~ country)      
```

## Using reorder() to arrange facets by average donation rate

```{r grouped_and_faceted-1, fig.width = 12, fig.height = 8}
p <- ggplot(data = organdata,
            mapping = aes(x = year, y = donors))
p + geom_line() + 
  facet_wrap(~ reorder(country, donors, na.rm = TRUE))      
```

- A trick for reversing the ordering: reorder on the negative of the `donors` variable:

```{r grouped_and_faceted-2, fig.width = 12, fig.height = 8}
p <- ggplot(data = organdata,
            mapping = aes(x = year, y = donors))
p + geom_line() + 
  facet_wrap(~ reorder(country, -donors, na.rm = TRUE))      
```



