---
title: "Tables and Text"
linktitle: "12: Tables and Text"
date: "2020-09-25"
menu:
  show:
    parent: Examples
    weight: 1
type: docs
bibliography: "../../static/bib/references.bib"
csl: "../../static/bib/chicago-fullnote-bibliography-no-bib.csl"
output:
  blogdown::html_page:
    toc: false
editor_options: 
  chunk_output_type: console
---

```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
``` 

```{r plot1, message = FALSE}
library(tidyverse)
library(socviz)
```

## Controlling jitter

```{r jitter}

p <- ggplot(data = organdata, 
            mapping = aes(x = donors,
                          y = reorder(country, donors, na.rm=TRUE),
                          color = world))

p + geom_jitter(position = position_jitter(height=0.15)) +
    labs(y = NULL) + 
    theme(legend.position = "top")

```

## Summarize better with dplyr

### 1.

```{r }
by_country <- organdata %>% 
              group_by(consent_law, country) %>%
              summarize(donors_mean = mean(donors, na.rm = TRUE),
                        donors_sd = sd(donors, na.rm = TRUE),
                        gdp = mean(gdp, na.rm = TRUE),
                        health = mean(health, na.rm = TRUE),
                        roads_mean = mean(roads, na.rm = TRUE),
                        cerebvas = mean(cerebvas, na.rm = TRUE))
                        
by_country                        
```


### 2. 

```{r }
by_country <- organdata %>% 
  group_by(consent_law, country) %>%
    summarize_if(is.numeric, 
                 list(~ mean(.x, na.rm = TRUE), 
                      ~ sd(.x, na.rm = TRUE))) %>%
    ungroup()

by_country
```


### 3. 

```{r }
by_country <- organdata %>% 
  group_by(consent_law, country) %>%
    summarize(across(where(is.numeric),
                     list(mean = mean, sd = sd),
                      na.rm = TRUE,
                      .names = "{col}_{fn}"),
              .groups = "drop")

by_country
```

## Free scales when faceting categorical variables

### No

```{r }
p <- ggplot(data = by_country,
            mapping = aes(x = donors_mean, 
                          y = reorder(country, 
                                      donors_mean)))

p + geom_point(size=3) +
    facet_wrap(~ consent_law) +
    labs(x = "Donor Procurement Rate",
         y = NULL)
```


## Closer but still no

```{r }
p <- ggplot(data = by_country,
            mapping = aes(x = donors_mean, 
                          y = reorder(country, 
                                      donors_mean)))


p + geom_point(size=3) +
    facet_wrap(~ consent_law, scales = "free_y") +
    labs(x= "Donor Procurement Rate",
         y= NULL)
```

## Good

```{r }
p <- ggplot(data = by_country,
            mapping = aes(x = donors_mean, 
                y = reorder(country, donors_mean)))

p + geom_point(size=3) +
    facet_wrap(~ consent_law, 
                 scales = "free_y", ncol = 1) +
    labs(x = "Donor Procurement Rate",
         y = NULL)
```

## There are a lot of variants in the point-and-range family

```{r }

p <- ggplot(data = by_country,
            mapping = aes(x = donors_mean,
                          y = reorder(country, donors_mean)))

p + geom_pointrange(mapping = aes(xmin = donors_mean - donors_sd, 
                                  xmax = donors_mean + donors_sd)) +
    labs(x = "Donor Procurement Rate", y = NULL)

```

## Plotting Text Directly

### As a rule, don't use geom_text(), use gg_repel()


```{r fig.cap="Popular and Electoral College Margins.",  out.width="100%", fig.width=12, fig.height=8, fig.fullwidth=TRUE}
library(ggrepel)

## Put long labels in named objects; good housekeeping

p_title <- "Presidential Elections: Popular & Electoral College Margins"
p_subtitle <- "1824-2016"
p_caption <- "Data for 2016 are provisional."
x_label <- "Winner's share of Popular Vote"
y_label <- "Winner's share of Electoral College Votes"

## Set a theme
theme_set(theme_minimal())

## Note use of geom_hline() and geom_vline()
p <- ggplot(data = elections_historic, 
            mapping = aes(x = popular_pct,
                          y = ec_pct,
                          label = winner_label))
                          
p + geom_hline(yintercept = 0.5, size = 1.4, color = "gray70") +
    geom_vline(xintercept = 0.5, size = 1.4, color = "gray70") +
    geom_point() +
    geom_text_repel() +
    scale_x_continuous(labels = scales::percent) +
    scale_y_continuous(labels = scales::percent) +
    labs(x = x_label,
         y = y_label,
         title = p_title,
         subtitle = p_subtitle,
         caption = p_caption)

```

- Control the height and width of your plot via `ggsave()` or in the options to your code chunk with `fig.width` and `fig.height`.


## Exploring covdata

```{r }

## Please reinstall covdata! It's been updated. Use at least version 0.5

# drat::addRepo("kjhealy")
# install.packages("covdata")

library(covdata)

```

### Short Term Morality Fluctuations

### Belgium (mortality rates, from `stmf`)

```{r }
stmf %>%
  filter(sex == "b", country_code == "BEL") %>%
  group_by(year, week) %>%
  mutate(yr_ind = year %in% 2020) %>%
  group_by(yr_ind, year, week) %>% 
  summarize(tot_death = sum(death_count), .groups = "drop") %>% 
  ggplot(aes(x = week, y = tot_death, color = yr_ind, group = year)) + 
  geom_line(size = 0.9) + 
  scale_color_manual(values = c("gray70", "red"), labels = c("2000-2019", "2020")) +
  labs(x = "Week of the Year", 
       y = "Total Deaths", 
       color = "Year",
       title = "Weekly recorded deaths in Belgium, 2000-2020") + 
  theme_minimal() + 
  theme(legend.position = "top")

```


### United States, (mortality counts, from `nchs_wdc`)

```{r }

nchs_wdc %>%
  filter(jurisdiction == "United States", cause == "All Cause", year > 2014) %>%
  group_by(year, week) %>%
  mutate(yr_ind = year %in% 2020) %>%
  filter(!(year == 2020 & week > 30)) %>%
  ggplot(aes(x = week, y = n, color = yr_ind, group = year)) +
  geom_line(size = 0.9) +
  scale_color_manual(values = c("gray70", "firebrick"), labels = c("2015-2019", "2020")) +
  scale_x_continuous(breaks = c(1, 10, 20, 30, 40, 50),
                     labels = as.character(c(1, 10, 20, 30, 40, 50))) +
  scale_y_continuous(labels = scales::comma) +
  labs(x = NULL,
       y = "Total Deaths",
       color = "Years",
       title = "Weekly recorded deaths from all causes in the United States",
       subtitle = "2020 data are for Weeks 1 to 30. Raw Counts.") +
  theme_minimal() +
  theme(legend.position = "top")


```
