---
title: "Problem set 1: Plotting Gapminder Data"
date: "2019-09-13"
due_date: "2019-09-20"
due_time: "6:00 PM"
editor_options: 
  chunk_output_type: console
---


## In this Problem Set, you will use R, RStudio, and ggplot to look at the `gapminder` data again. 

## Our goal is to create an RStudio project, load the required libraries, and begin looking at our data. 

##  **Create** a new RStudio project and place it on your computer somewhere. 

## Open that new folder in Windows File Explorer or macOS Finder (however you navigate around the files on your computer), and create a folder there named `figures` 

## **Create** a new R Markdown file and save it in your project. 

## In RStudio go to File > New File > R Markdown…, choose the default options, and delete all the placeholder text in the new file except for the metadata at the top, which is between `---` and `---`. 


## **Verify** that your project folder is structured like this:

    ```
    your-project-name/
      your-analysis.Rmd
      your-project-name.Rproj
      figures/
        <EMPTY>
    ```

## **Load** the libraries required for the analysis

## Create an R code chunk that runs this code:

```{r setup}

library(tidyverse)
library(gapminder)
library(socviz)

```

## **Examine** the Gapminder Data

```{r gapminder}
gapminder
```


## **Create** the following plots (Make a new code chunk for each plot)

1. A scatter plot (using `geom_point()`) of GDP per Capita (on the x-axis) and Life Expectancy (on the y-axis). 
2. Adjust the plot so that the x-axis scale is in log units rather than raw GDP, and properly label the axes. 
3. Using the `subset()` function, choose a *single year* and make a histogram of country populations. *Hint:* Use `==` and not `=` when asking `subset` to pick a particular value of `year`.
4. Experiment with the `binwidth` argument (or alternatively the `bins` argument) as the note from `stat_bin()` suggests.
5. Try adding a log scale to the x-axis
6. Facet the plot by continent
7. Subset the data so that you pick out the following four years: 1952, 1972, 1992, and 2002. Plot a histogram of life expectancy faceted by year. 

## Briefly answer the following questions about this final plot:

- What bin width did you pick for the plot? Why
- What do the histograms appear to show about life expectancy? 


# Finish

Knit the completed R Markdown file as a Word or PDF document (use the "Knit" button at the top of the script editor window). Save it with a name of the form `lastname_firstname_ps01` and upload it to the Sakai dropbox.
