---
title: "Problem set 3: Dog Names and People Names"
date: "2019-10-11"
due_date: "2019-10-18"
due_time: "6:00 PM"
editor_options: 
  chunk_output_type: console
---


## In this Problem Set, we'll be looking at the Names of New York Dogs and American Humans

## If you haven't done so already, download the skeleton project that contains the NYC dogs data

You can clone it in GitHub if you're familiar with that process: <https://github.com/kjhealy/nyc_tmp>

Alternatively, download a zip file of the project: <https://github.com/kjhealy/nyc_tmp/archive/master.zip>

When you download this file, *make sure you unzip it* rather than just looking inside it with your file manager. 

## Open the project in R Studio by double-clicking the `nyc_tmp.Rproj` file. **Verify** (in R Studio's file viewer pane) that your project folder is structured like this:

    ```
    nyc_tmp/
      data/
        nyc_bites.rda
        nyc_license.rda
        nyc_zips.rda
      .gitignore
      .here
      dogs.Rmd
      nyc_tmp.Rproj
        
    ```

## Open the `dogs.Rmd` file and do the analysis in that.

## Load the required libraries

```{r setup}

library(tidyverse)
library(stringr)
library(socviz)
library(babynames)
library(sf)

```

## Load the dogs data

If you copy this code chunk into your working file, remove `eval = FALSE` from them, otherwise it won't run. 

```{r eval = FALSE}

load("data/nyc_license.rda")
load("data/nyc_zips.rda")
load("data/nyc_bites.rda")

```



## Take a look at the babynames data

```r{}

babynames

```

## Remember how we've been working with the NYC dogs data

Reproduce the code below in your own analysis. 

```{r eval = FALSE}

boro_names <- c("Manhattan", "Queens", 
                "Brooklyn", "Bronx", 
                "Staten Island")
no_name <- c("Unknown", "Name Not Provided")

top_names <- nyc_license %>% 
  filter(borough %in% boro_names, 
         animal_name %nin% no_name) %>% 
  group_by(borough, animal_name) %>%
  tally() %>%
  mutate(freq = n / sum(n), 
         pct = freq * 100) %>% 
  top_n(5) %>%
  arrange(desc(pct)) 
  
p <- ggplot(data = top_names,
            mapping = aes(x = reorder(animal_name, pct), 
                          y = pct))

p + geom_point(size = 3) + 
  coord_flip() + 
  facet_wrap(~ reorder(borough, -pct), 
             scales = "free_y", ncol = 1) + 
  labs(x = NULL, 
       y = "Percent", 
       title = "Most Popular Dog Names") + 
  theme_minimal()
```


## Questions to answer

1. Briefly explain in your own words what the sequence of operations from `nyc_license %>%` to `arrange(desc(pct))` are doing to the data in the code chunk listed above.
2. Produce a table and accompanying figure that shows the five most popular names for male and female dogs, instead of for each borough. 
3. Produce a table and accompanying figure that shows the five most popular names for male and female dogs, *within* each borough. 
4. Produce a map showing where dogs named Max are most commonly found in New York City.
6. Now look at the `babynames` dataset. What were the five most popular names for girls in 1900? What where they in 2000?
7. Make a figure tracing the popularity of any two of the following names: Shirley, Linda, Bittany, Emma, Ella, Mark, Oliver, Brayden, Michael. (Remember to filter the data by sex before trying to draw any plots.)
8. Place the two names on the same plot.
9. Can you think of a way to get a sense of the relative heterogeneity of boy vs girl names over time?


# Finish

Knit the completed R Markdown file as a Word or PDF document (use the "Knit" button at the top of the script editor window). Save it with a name of the form `lastname_firstname_ps02` and upload it to the Sakai dropbox.
