---
title: "Vision, Data, and Design"
linktitle: "05: Vision, Data, and Design"
date: "2020-08-30"
due_date: "2020-09-04"
due_time: "11:59 AM"
output:
  blogdown::html_page:
    toc: false
menu:
  work:
    parent: Work
    weight: 1
type: docs
weight: 1
editor_options: 
  chunk_output_type: console
---

## Problem Set 2

1. Use RStudio to create a file named something like `lastname_firstname_ps02.Rmd`
2. Set up your file with a code chunk that loads the `tidyverse`, and `gapminder` packages:

```{r plot1, message = FALSE}
library(tidyverse)
library(gapminder)
library(socviz)
```

3. Then use R to draw these plots:

```{r plot-samples-1, fig.cap = "GDP and Life Expecancy 1", echo = FALSE, message = FALSE}

ggplot(data = gapminder,  
     mapping = aes(x = gdpPercap, y = lifeExp)) + 
     geom_point() + 
     geom_smooth() + 
     labs(x = "GDP Per Capita",
     y = "Life Expectancy")

```


```{r plot-samples-2, fig.cap = "GDP and Life Expecancy 2", echo = FALSE, message = FALSE}

ggplot(data = gapminder,  
     mapping = aes(x = gdpPercap, y = lifeExp)) + 
     geom_point() + 
     geom_smooth() + 
     scale_x_log10(labels = scales::dollar) + 
     labs(x = "GDP Per Capita (log scale)",
     y = "Life Expectancy")

```

```{r plot-samples-3, fig.cap = "GDP and Life Expecancy 3", echo = FALSE, message = FALSE}
ggplot(data = gapminder,  
     mapping = aes(x = gdpPercap, y = lifeExp, size = pop)) + 
     geom_point() + 
     geom_smooth(method = "lm") + 
     scale_x_log10(labels = scales::dollar) + 
     labs(x = "GDP Per Capita (log scale)",
     y = "Life Expectancy", size = "Population")

```

```{r plot-samples-4, fig.cap = "GDP and Life Expectancy 4", echo = FALSE, message = FALSE}
ggplot(data = gapminder,  
     mapping = aes(y = gdpPercap, x = lifeExp, size = pop)) + 
     geom_point(color = "gray20", alpha = 0.3) + 
     geom_smooth() + 
     scale_y_log10(labels = scales::dollar) + 
     labs(y = "GDP Per Capita (log scale)",
     x = "Life Expectancy", size = "Population")

```

4. Discuss in a few words 
- a) What the mappings are in each plot.
- b) How the plots do or do not illustrate what Tufte, Wilke, or Ware discuss in the readings.
- c) The differences between Figures 2, 3, and 4, and what difference if any those differences make to your interpretation of the figure.
