---
title: "Notes for November 6th/8th: Geom Raster"
date: "2019-10-27"
citeproc: false
bibliography: ../../static/bib/references.bib
csl: ../../static/bib/chicago-syllabus-no-bib.csl
output:
  blogdown::html_page:
    template: ../../pandoc/toc-title_html.template
    toc: true
editor_options: 
  chunk_output_type: console
---

# `geom_tile()` and `geom_raster()`

First let's set up some libraries and load some data. 

```{r libs}
library(tidyverse)
library(socviz)
library(demog)
```

Here's the birth rate data:

```{r }
okboomer
```

We can make a time series or trend plot of this data with `geom_line()`:

```{r, fig.fullwidth = TRUE, fig.height = 3, fig.width = 8}

okboomer %>%
    filter(country == "United States") %>%
    ggplot(aes(x = date, y = births_pct_day)) +
    geom_line(size = 0.9) +
    labs(x = "Year",
         y = "Average daily births per million") 

```

But we can also use `geom_tile()` to create an interesting plot. To get it right, we will convert years and months to _factors_ or categorical variables, like this:

```{r }
okboomer <- okboomer %>%
    mutate(year_fct = factor(year,  
                             levels = rev(unique(year)), 
                             ordered = TRUE),
           month_fct = factor(month,
                              levels = c(1:12),
                              labels = c("Jan", "Feb", "Mar", "Apr",
                                    "May", "Jun", "Jul", "Aug",
                                    "Sep", "Oct", "Nov", "Dec"),
                              ordered = TRUE)) %>%
    select(year, month, 
           year_fct, month_fct, everything())

```

We choose the levels and ordering of the categories considering the plot we want to make. In this case we'll make the plot tall rather than wide by putting year on the y axis. 

```{r, fig.fullwidth = TRUE, fig.height = 20, fig.width = 6}

okboomer %>%
    filter(country == "United States") %>%
    ggplot(aes(x = month_fct, y = year_fct)) +
    geom_tile(mapping = aes(fill = births_pct_day), 
              color = "white") + 
    scale_x_discrete(position = "top") +              
    scale_y_discrete(breaks = seq(1940, 2010, 5)) +    
    scale_fill_viridis_c(option = "B") + 
    labs(x = NULL, y = NULL, fill = NULL, title = "Monthly Birth Rates",
         subtitle = "Average births per million people per day.",
         caption = "Data: US Census Bureau.")
 
```

