---
title: "Notes for November 13th/15th: Networks"
date: "2019-10-27"
citeproc: false
bibliography: ../../static/bib/references.bib
csl: ../../static/bib/chicago-syllabus-no-bib.csl
output:
  blogdown::html_page:
    template: ../../pandoc/toc-title_html.template
    toc: true
editor_options: 
  chunk_output_type: console
---

# Network data in ggplot

First let's set up some libraries. 

```{r libs}
library(tidyverse)
library(socviz)

## Install if necessary with install.packages()
library(ggraph)
library(tidygraph)
library(graphlayouts)
```

Next, load the data package. If you don't have the `kjhnet` package installed, then use `drat` to tell R where to find it adn then `install.packages()`

```{r data}

# drat::addRepo("kjhealy")
# install.packages("kjhnet")

library(kjhnet)

```

```{r theme, echo = FALSE, message = FALSE, warning = FALSE}
library(showtext)
showtext_auto()

library(myriad)
import_myriad_semi()

theme_set(theme_myriad_semi())
set_graph_style(family = "Myriad Pro SemiCondensed")

```

## Paul Revere


First the groups data:

```{r persons-r}
revere_groups
```

We'll calculate a centrality measure (that we might look at separately) and then plot the edges weighted by the number of ties between organzations.

```{r groups-plot}
revere_groups %>%
  mutate(centrality = centrality_degree()) %>%
  ggraph(layout = "kk") +
  geom_edge_link(aes(width = value), color = "gray80") +
  geom_node_label(aes(label = name))
```

Next, the persons data:

```{r persons-dat}
revere_persons
```

```{r persons-plot}
revere_persons %>%
  activate(nodes) %>%
  mutate(centrality = centrality_eigen()) %>%
  ggraph(layout = "stress") +
  geom_edge_link0(aes(filter = value > 1, width = value), color = "gray90") +
  geom_node_point() +
  geom_node_label(aes(filter = centrality > 0.9, label = name), size = rel(2.5))
```

## The Iliad Data

These data are from Rossman (2017), see <https://osf.io/jasf4/> for details.

```{r iliad}
il_tidy
```

Here's the overall network.

```{r iliad-plot1}
il_tidy %>%
  activate(nodes) %>%
  ggraph(layout = "fr") +
  geom_edge_link(color = "gray80") +
  geom_node_point(aes(color = affil)) +
  scale_color_manual(values = c("blue", "red"), labels = c("Athenian", "Trojan")) +
  guides(color = guide_legend(title = "Side", )) +
  labs(title = "Killings in The Iliad") +
  theme(plot.title = element_text(size = rel(3)))
```

Calculate and plot the Betweenness Centrality of the male characters. 

```{r iliad-between}
il_tidy %>%
  activate(nodes) %>%
  mutate(centrality = centrality_degree(),
         betweenness = centrality_betweenness()) %>%
  arrange(desc(betweenness)) %>%
    top_n(10, wt = betweenness) %>%
    as_tibble() %>%
  ggplot(mapping = aes(x = betweenness, y = reorder(name, betweenness), color = affil)) + 
  geom_point(size = 3) + 
  labs(x = "Betweenness", y = NULL, color = "Side") +
  theme_minimal() + 
  theme(legend.position = "top")
```

And the "Google centrality" (i.e., Pagerank) of the male characters. Notice that we calculate this on the edges, not the nodes.

```{r iliad-pagerank}
il_tidy %>%
    activate(edges) %>%
  filter(act == "kills") %>%
  reroute(from = to, to = from) %>%
  activate(nodes) %>%
  mutate(p_rank = centrality_pagerank()) %>%
   top_n(10, wt = p_rank) %>%
   as_tibble() %>%
  ggplot(mapping = aes(x = p_rank,
                       y = reorder(name, p_rank),
                       color = affil)) +
  geom_point(size = 3) +
    labs(x = "PageRank Centrality", y = NULL,
         color = "Side") +
  theme_minimal() +
  theme(legend.position = "top")
```

Finally a graph with labels. You may need to install the `prismatic` library here. 

```{r iliad-labeled, fig.width = 10, fig.height = 10, fig.fullwidth = TRUE}
label_colors <- c(prismatic::clr_lighten("blue", 0.7), 
                  prismatic::clr_lighten("red", 0.7))
il_tidy %>%
    activate(nodes) %>%
    mutate(centrality = centrality_degree(mode = "out")) %>%
  ggraph(layout = 'graphopt') +
  geom_edge_link(aes(start_cap = label_rect(node1.name),
                     end_cap = label_rect(node2.name)),
                 arrow = arrow(length = unit(1.5, 'mm'))) +
  geom_node_point(aes(color = affil)) +
  scale_color_manual(values = c("blue", "red"), labels = c("Athenian", "Trojan")) +
  guides(color = FALSE, fill = FALSE) +
  geom_node_label(aes(filter = centrality > 3, label = name, fill = affil), size = rel(2.5)) +
  scale_fill_manual(values = label_colors) + 
  labs(title = "Killings in The Iliad") +
  theme(plot.title = element_text(size = rel(3)))

```
