---
title: "Problem set 2: MacArthur Award Data"
date: "2019-09-27"
due_date: "2019-10-04"
due_time: "6:00 PM"
editor_options: 
  chunk_output_type: console
---


## In this Problem Set, you will use R, RStudio, and ggplot to look at a freshly-collected data and not yet cleaned dataset about MacArthur Award winners.


##  **Create** a new RStudio project and place it on your computer somewhere. 

## Open that new folder in Windows File Explorer or macOS Finder (however you navigate around the files on your computer), and create a folder there named `figures` 

## **Create** a new R Markdown file and save it in your project. 

## In RStudio go to File > New File > R Markdown…, choose the default options, and delete all the placeholder text in the new file except for the metadata at the top, which is between `---` and `---`. 


## **Verify** that your project folder is structured like this:

    ```
    your-project-name/
      your-analysis.Rmd
      your-project-name.Rproj
      data/
        <EMPTY>
      figures/
        <EMPTY>
    ```

## **Load** the libraries required for the analysis

```{r setup}

library(tidyverse)
library(socviz)

```

## **Download** the MacArthur Data

The data are at the following URL: 

`https://visualizingsociety.com/data/macarthur.csv`

Use the `read_csv()` function to get this data into an object called `macarthur`. You can do this more than one way. Either grab the data directly from the url, or save it to your project's `data/` folder first and then load it.

## If you download the data directly, **save** the data file to your computer, in the `data/` folder

Use `write_csv()` to do this.

## **Examine** the MacArthur Data

```{r gapminder, eval = FALSE}
macarthur
```

The variables in the data are as follows:

* `name`: The name of the awardee
* `url`: Their Wikipedia page, if they have one
* `year`: The year of their award
* `body_out`: Their birth year as scraped from their Wikipedia page
* `bday_out`: Their birth year as extracted from their Wikipedia bio card
* `born`: Our best first guess as to the awardee's birth year. (It is the value of `bday_out` if there is one, otherwise the value of `body_out`, if there is one.)
* `born_rc`: Our best guess of as to the awardee's birth year, with some manual changes applied
* `age`: Our calculation of the awardee's age at the time of their award. This is `year` minus `born_rc`.


## Questions to answer

1. How much missing data is there in this file, and what is missing? (Hint: missing data is designated by `NA`. The `is.na()` function returns `TRUE` if an observation is missing and `FALSE` if it is not.)
2. Let's say we are interested in whether the age profile of MacArthur award recipients has changed over time. Draw some plots to help you investigate this question. 
3. Can we use `geom_boxplot()` to look at the age distribution of awardees within each year? (Hint: try using the `socviz` library's `int_to_year()` function to change the _year_ variable from an integer to a date. Alternatively, try using ``factor()` to make `year` a categorical variable.)
4. Does the missing data show any structure that you can see, or is it missing at random?
5. Can you find any errors in the data? What sort of strategies might we employ for finding errors?
6. Can you make a plot that picks out or highlights our very own Professor Tung in this data?

# Finish

Knit the completed R Markdown file as a Word or PDF document (use the "Knit" button at the top of the script editor window). Save it with a name of the form `lastname_firstname_ps02` and upload it to the Sakai dropbox.
