---
title: "Problem set 5: Birth and Death Rates"
date: "2019-11-08"
due_date: "2019-11-15"
due_time: "6:00 PM"
editor_options: 
  chunk_output_type: console
---


## In this Problem Set, we will examine birth and death rates for various countries

The data is available as an R package. To install it, do the following.

If you haven't already, install and load `drat`:

1. Install the `drat` package with `install.packages(drat)`
2. Load it with `library(drat)`
3. Add the repository where the data is: `drat::addRepo("kjhealy")`

You can now install `demog` with

4. `install.packages("demog")`

Note that the `demog` package has been updated since last week! So `install.packages("demog")` even if you've done it before.

## Create a project for the assignment

## Open the project in RStudio and make an Rmd file for the analysis called something like `okboomer.Rmd`

## Load the required libraries

```{r setup}

library(tidyverse)
library(socviz)
library(demog)

```

## Take a look at the data

There are various data files included in this package. You can get a brief summary of each variable in the dataset by looking at the Help file in RStudio for the `demog` package, or by looking at the documentation on the package homepage: <http://kjhealy.github.io/demog>.

Start with birth rates in the US and England/Wales:

```{r}

okboomer

```

## Questions to answer on the Birth Rates data

1. What is the unit of observation in this dataset?
2. Make a plot of birth rates for the United States.
3. Over the years, which _month_ has the highest average number of births per capita per day?
4. Draw a plot showing the seasonality of births in the United States for each decade in the dataset. _Hint:_ Consider using `coord_polar()` to draw a circular chart.


## Mortality rates

The mortality rate data is in a nested tibble:

```{r }
mortality
```

The `data` column of this tibble contains each country's estimated male, female, and total morality rates for each age-year combination. 

## Questions to answer on the Mortality data

Choose a country from the data to focus on. 

1. Plot the combined (`total`) mortality rate at age 0, age 5, age 50, and age 70 across all years in your country's data. 
2. Create a Lexis Surface plot or heatmap for that country's male mortality rate. _Hint:_ Use `geom_raster()` or `geom_tile()` to graph the mortality rate mapped to the `fill` aesthetic.
3. Calculate the _ratio_ of male to female mortality rates for the country you've chosen, and the _percentage difference_ between male and female mortality rates. Make a plot of the results. What sort of color scale should you use for measures like these, and why? Briefly discuss any patterns you can discern in the data you graph.  

# Finish

Knit the completed R Markdown file as a Word or PDF document (use the "Knit" button at the top of the script editor window). Save it with a name of the form `lastname_firstname_ps05` and upload it to the Sakai dropbox.
